---
title: "V5.Rmd"
author: "Sameer"
date: "2025-04-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(smooth)
library(zoo)
library(kableExtra)
```

# Data Wrangling

Transform the hourly data into daily data:

Cleaned Dataframes: daily_load, daily temp, daily_humidity

```{r}
# Load the demand excel file
load_df <- read_excel("../Data/load.xlsx")
  
daily_load <- load_df %>% 
  group_by(meter_id, date) %>%
  summarize(
    daily_avg = mean(c_across(1:24), na.rm = TRUE),
    .groups   = "drop"
  )

# Load the temperature file
temp_df <- read_excel("../Data/temperature.xlsx")

daily_temp <- temp_df[1:3] %>% 
  group_by(date) %>%
  summarize(
    daily_avg = mean(t_ws1, na.rm = TRUE),
    .groups   = "drop"
  )

# Load the humidity file
humidity_df <- read_excel("../Data/relative_humidity.xlsx")

daily_humidity <- humidity_df[1:3] %>% 
  group_by(date) %>%
  summarize(
    daily_avg = mean(rh_ws1, na.rm = TRUE),
    .groups   = "drop"
  )

daily_temp_for <- read.csv("../Data/Temp_Forecast.csv") 
temp_for <- msts(daily_temp_for$x, seasonal.periods =c(7,365.25), start=c(2011,1,1))

daily_humidity_for <- read.csv("../Data/Humidity_Forecast.csv") 
humidity_for <- msts(daily_humidity_for$x, seasonal.periods =c(7,365.25), start=c(2011,1,1))
```

```{r}
TS_Plot <- 
  ggplot(daily_load, aes(x=date, y=daily_avg)) +
      geom_line()
plot(TS_Plot)

#ACF and PACF plots
par(mfrow=c(1,2))
ACF_Plot <- Acf(daily_load$daily_avg, lag = 40, plot = TRUE)
PACF_Plot <- Pacf(daily_load$daily_avg, lag = 40)
par(mfrow=c(1,1))
```

ACF plot shows slow decay, which is a sign of non-stationarity. Proceed with differencing to test this.

```{r}
diff_load_series <- diff(daily_load$daily_avg, differences = 1)
plot(diff_load_series, type = "l", main = "Differenced Series")
acf(diff_load_series, main = "ACF of Differenced Series")
pacf(diff_load_series, main = "PACF of Differenced Series")

```


```{r}
# Create time series for differenced load series
load_ts <- msts(daily_load$daily_avg, 
                seasonal.periods =c(7,365.25), 
                start=c(2005,1,1))

diff_load_ts <- msts(diff_load_series, 
                seasonal.periods =c(7,365.25), 
                start=c(2005,1,1))

# Create time series for temperature 
temp_ts <- msts(daily_temp$daily_avg, 
                seasonal.periods =c(7,365.25), 
                start=c(2005,1,1))

# Create time series for humidity
humidity_ts <- msts(daily_humidity$daily_avg, 
                seasonal.periods =c(7,365.25), 
                start=c(2005,1,1))
```

## Decomposing time series objects
```{r}
load_ts %>% mstl() %>%
  autoplot()

diff_load_ts %>% mstl() %>%
  autoplot()

temp_ts %>% mstl() %>%
  autoplot()

humidity_ts %>% mstl() %>%
  autoplot()
```

## Creating test and training datasets 
```{r}
#create a subset for training purpose
n_for = 365
ts_diff_load_train <- subset(diff_load_ts,
                        end = length(diff_load_ts)-n_for)

#create a subset for testing purpose
load_test <- tail(load_ts, 365)

ts_temp_train <- subset(temp_ts,
                        end = length(temp_ts)-n_for)

ts_temp_test <- subset(temp_ts,
                       start = length(temp_ts)-n_for)

ts_humidity_train <- subset(humidity_ts,
                        end = length(humidity_ts)-n_for)

ts_humidity_test <- subset(humidity_ts,
                       start = length(humidity_ts)-n_for)

```

### Model 1: STL + ETS

```{r ETS, echo=TRUE, message=FALSE, warning=FALSE}
#Fit and forecast STL + ETS model to data
ETS_fit <-  stlf(ts_diff_load_train,h=365)

#Plot foresting results
autoplot(ETS_fit) + ylab("Load")

#Plot model + observed data
autoplot(load_ts) +
  autolayer(ETS_fit, series="STL + ETS",PI=FALSE) +
  ylab("Load")+
  theme_minimal()

# Reconstruct forecasted levels
# 1. Get the last observed value from the ORIGINAL training data (before differencing)
last_observed <- tail(load_ts, n = length(ts_diff_load_train) + 1)[length(ts_diff_load_train)]

# 2. Get the forecasted differenced values from your model
STL_forecast_diff <- ETS_fit$mean

# 3. Undifference: Reconstruct the level forecast
STL_forecast_levels <- cumsum(c(last_observed, STL_forecast_diff))[-1]

# 4. Now compare to the original (non-differenced) test set
accuracy(STL_forecast_levels, load_test)


```

### Model 2: ARIMA + FOURIER terms

```{r ARIMA, echo=TRUE, message=FALSE, warning=FALSE}
ARIMA_Four_fit <- auto.arima(ts_diff_load_train, 
                             seasonal=FALSE, 
                             lambda=0,
                             xreg=fourier(ts_diff_load_train, 
                                          K=c(2,6))
                             )

#Forecast with ARIMA fit
#also need to specify h for fourier terms
ARIMA_Four_for <- forecast(ARIMA_Four_fit,
                           xreg=fourier(ts_diff_load_train,
                                        K=c(2,6),
                                        h=365),
                           h=365
                           ) 

#Plot forecasting results
autoplot(ARIMA_Four_for) + ylab("Load")

#Plot model + observed data
autoplot(load_ts) +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Load")+
  theme_minimal()
#Plot forecasting results
autoplot(ARIMA_Four_for) + ylab("Load")

#Plot model + observed data
autoplot(load_ts) +
  autolayer(ARIMA_Four_for, series="ARIMA_FOURIER",PI=FALSE) +
  ylab("Load")+
  theme_minimal()

# Accuracy test using the direct forecast
ARIMA_scores <- accuracy(ARIMA_Four_for$mean, load_test)
ARIMA_scores
```

### Model 3: ARIMA + Fourier (with temperature and humidity)
```{r}
# Combine Fourier terms, temperature, and humidity for training
ts_temp_train_subset <- ts_temp_train[1:length(ts_diff_load_train)]
ts_humidity_train_subset <- ts_humidity_train[1:length(ts_diff_load_train)]

# Combine Fourier terms, temperature, and humidity for training
train_regressors <- cbind(fourier(ts_diff_load_train, K=c(2,6)),
                          ts_temp_train_subset,
                          ts_humidity_train_subset)

ARIMA_Env_fit <- auto.arima(ts_diff_load_train,
                             seasonal=FALSE,
                             lambda=0,
                             xreg=train_regressors)

ts_temp_test <- ts_temp_test[1:length(load_test)]
ts_humidity_test <- ts_humidity_test[1:length(load_test)]

# Create future regressors for the forecast period
future_regressors <- cbind(fourier(ts_diff_load_train, K=c(2,6), h=length(load_test)),
                           ts_temp_test,
                           ts_humidity_test)

# Forecast with ARIMA fit, including future regressors
ARIMA_Env_for <- forecast(ARIMA_Env_fit,
                           xreg=future_regressors,
                           h=length(load_test))

# Plot forecasting results
autoplot(ARIMA_Env_for) + ylab("Load")

# Plot model + observed data
autoplot(load_ts) +
  autolayer(ARIMA_Env_for, series="ARIMA_ENV", PI=FALSE) +
  ylab("Load")+
  theme_minimal()

# Accuracy test against load_test
ARIMA_Env_scores <- accuracy(ARIMA_Env_for$mean, load_test)
ARIMA_Env_scores
```

### Model 4: Neural Network Time Series Forecasts

```{r NNETAR, echo=TRUE, message=FALSE, warning=FALSE}
#You can play with the different values for p and P, you can also use xreg with Fourier term to model the multiple seasonality

#NN_fit <- nnetar(ts_act_power_daily_train,p=1,P=1)
NN_fit <- nnetar(ts_diff_load_train,
                 p=7,
                 P=1,
                 xreg=fourier(ts_diff_load_train, K=c(2,4)))

#NN_for <- forecast(NN_fit, h=365) 
NN_for <- forecast(NN_fit, h=365,xreg=fourier(ts_diff_load_train, 
                                          K=c(2,4),h=365))

#Plot foresting results
autoplot(NN_for) +
  ylab("Load") 

#Plot model + observed data
autoplot(load_ts) +
  autolayer(NN_for, series="Neural Network",PI=FALSE)+
  ylab("Load") 

#Accuracy check
NN_scores <- accuracy(NN_for$mean,load_test)
NN_scores
```

### Model 5: Neural Network Time Series Forecasts with temperature and humidity

```{r NNETAR, echo=TRUE, message=FALSE, warning=FALSE}
train_regressors <- cbind(fourier(ts_diff_load_train, K=c(2,4)),
                          ts_temp_train_subset,
                          ts_humidity_train_subset)

# Create future regressors for the forecast period
future_regressors_nn <- cbind(fourier(ts_diff_load_train, K=c(2,4), h=length(load_test)),
                           ts_temp_test,
                           ts_humidity_test)

#NN_fit <- nnetar(ts_act_power_daily_train,p=1,P=1)
NN_w_temp_humid_fit <- nnetar(ts_diff_load_train,
                 p=7,
                 P=1,
                 xreg=fourier(ts_diff_load_train, K=c(2,4)))

#NN_for <- forecast(NN_fit, h=365) 
NN_w_temp_humid_for <- forecast(NN__w_temp_humid_fit, h=365,xreg=fourier(ts_diff_load_train, 
                                          K=c(2,4),h=365))

#Plot foresting results
autoplot(NN_w_temp_humid_for) +
  ylab("Load") 

#Plot model + observed data
autoplot(load_ts) +
  autolayer(NN_w_temp_humid_for, series="Neural Network",PI=FALSE)+
  ylab("Load") 

#Accuracy check
NN_w_temp_humid_scores <- accuracy(NN_w_temp_humid_for$mean, load_test)
NN_w_temp_humid_scores
```

Forecast and print out results from Model 5
```{r}
h_future <- 59

NN_w_temp_humid_fit <- nnetar(load_ts,
                 p=7,
                 P=1,
                 xreg=fourier(load_ts, K=c(2,4)))

# Forecast for h_future days (2011)
NN_w_temp_humid_for <- forecast(NN_w_temp_humid_fit, h=h_future,xreg=fourier(load_ts, 
                                          K=c(2,4),h=h_future))

NN_result <- as.data.frame(NN_w_temp_humid_for$mean)
print(NN_result)
```

