---
title: "V6"
author: "Sameer"
date: "2025-04-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(forecast)  
library(Kendall)
library(tseries)
library(smooth)
library(zoo)
library(kableExtra)
```

# Data Wrangling

Transform the hourly data into daily data:

Cleaned Dataframes: daily_load, daily temp, daily_humidity

```{r}
# Load the demand excel file
load_df <- read_excel("../Data/load.xlsx")
  
daily_load <- load_df %>% 
  group_by(meter_id, date) %>%
  summarize(
    daily_load_avg = mean(c_across(1:24), na.rm = TRUE),
    .groups   = "drop"
  )

# Load the temperature file
temp_df <- read_excel("../Data/temperature.xlsx")

daily_temp <- temp_df[1:3] %>% 
  group_by(date) %>%
  summarize(
    daily_temp_avg = mean(t_ws1, na.rm = TRUE),
    .groups   = "drop"
  )

# Load the humidity file
humidity_df <- read_excel("../Data/relative_humidity.xlsx")

daily_humidity <- humidity_df[1:3] %>% 
  group_by(date) %>%
  summarize(
    daily_hum_avg = mean(rh_ws1, na.rm = TRUE),
    .groups   = "drop"
  )
```

Next, we plot temperature and humidity against load
Clear trend with temperature; not so much with humidity
Proceed with only temperature
```{r}
daily_all_avg <- daily_load %>%
                 left_join(daily_temp, by = "date") %>%
                 left_join(daily_humidity, by = "date")

#plot the dataset
ggplot(data = daily_all_avg, aes(x=daily_temp_avg, y = daily_load_avg)) +
              geom_point() +
              geom_smooth(method = lm)

ggplot(data = daily_all_avg, aes(x=daily_hum_avg, y = daily_load_avg)) +
              geom_point() +
              geom_smooth(method = lm)
```

Plot load and temperature individually to determine seasonality
```{r}
load_plot <- 
  ggplot(daily_load, aes(x=date, y=daily_load_avg)) +
      geom_line()
plot(load_plot)

temp_plot <- 
  ggplot(daily_temp, aes(x=date, y=daily_temp_avg)) +
      geom_line()
plot(temp_plot)
```

```{r}
# Create time series for load series
load_ts <- msts(daily_load$daily_load_avg, 
                seasonal.periods =c(7,365.25), 
                start=c(2005,1,1))

# Create time series for temperature 
#temp_ts <- msts(daily_temp$daily_temp_avg, 
                #seasonal.periods =c(7,365.25), 
                #start=c(2005,1,1))

temp_ts <- ts(daily_temp$daily_temp_avg,
              frequency = 365.25, # Approximate frequency for yearly data
              start = c(2005, 1))
```

Look at ACF plots for load and temp
```{r}
ACF_load_Plot <- Acf(daily_load$daily_load_avg, lag = 40, plot = TRUE)
ACF_temp_Plot <- Acf(daily_temp$daily_temp_avg, lag = 40)
```


## Decomposing time series objects
```{r}
load_ts %>% mstl() %>%
  autoplot()

temp_ts %>% mstl() %>%
  autoplot()
```

## Creating test and training datasets 
```{r}
#create a subset for training purpose
n_for = 365

ts_load_train <- subset(load_ts,
                        end = length(load_ts)-n_for)
#create a subset for testing purpose
ts_load_test <- subset(load_ts,
                        start = length(load_ts)-n_for)

ts_temp_train <- subset(temp_ts,
                        end = length(temp_ts)-n_for)

ts_temp_test <- subset(temp_ts,
                       start = length(temp_ts)-n_for)

#ts_temp_test <- ts_temp_test[1:length(load_test)]
```

#Train a NN model without temperature
```{r}
K = c(2,4)
for_period = 59
NN_forecast <- function(ts, p, P, K) {
  model_fit <- nnetar(ts,
                      p = p,
                      P = P,
                      xreg = fourier(ts, K = K))
  model_forecast <- forecast(model_fit, h=59, xreg=fourier(ts, K=K, h=for_period))
}
forecast_1 <- NN_forecast(load_ts, 1, 1, K)
NN_result_1 <- as.data.frame(forecast_1$mean)
```


```{r}
#Forecast future temperature values
h_future <- 59 
K_temp <- 2

#Fit NN model to forecast temp
temp_fit_fourier <- nnetar(temp_ts,
                           p = 7,
                           P = 1,
                           xreg = fourier(temp_ts, K = K_temp))

# Or using ARIMA:
temp_arima_fit_fourier <- auto.arima(temp_ts,
                                     xreg = fourier(temp_ts, K = K_temp))

# Generate future Fourier terms for temperature (based on the historical temperature series)
fourier_future_temp <- fourier(temp_ts, K = K_temp, h = h_future)

# Forecast temperature using nnetar
temp_forecast_fourier_nn <- forecast(temp_fit_fourier,
                                     h = h_future,
                                     xreg = fourier_future_temp)

# Forecast temperature using ARIMA
temp_forecast_fourier_arima <- forecast(temp_arima_fit_fourier,
                                        h = h_future,
                                        xreg = fourier_future_temp)

# The forecasted temperature series will be in:
# temp_forecast_fourier_nn$mean
# temp_forecast_fourier_arima$mean
```

Train a NN model with only temperature
```{r NNETAR, echo=TRUE, message=FALSE, warning=FALSE}
h_future <- 59
K=c(2,4)

#Train fourier model on Load
fourier_train <- fourier(ts_load_train, K = K)
#Bind temperature training daatset
train_regressors_with_temp <- cbind(fourier_train, ts_temp_train)
#Fit NN model on training data of load and temperature
NN_fit_with_temp_fourier <- nnetar(ts_load_train,
                                   p = 7,
                                   P = 1,
                                   xreg = train_regressors_with_temp)

#Do the same but for forecasted load and temperature
fourier_future <- fourier(load_ts, K = K, h = h_future)
future_regressors_with_temp <- cbind(fourier_future, temp_forecast_fourier_nn$mean)

NN_for_with_temp_fourier <- forecast(NN_fit_with_temp_fourier,
                                      h = h_future,
                                      xreg = future_regressors_with_temp)

#Look at accuracy scores
NN_scores <- accuracy(NN_for_with_temp_fourier$mean, ts_load_test)
NN_scores

#Print out results
NN_result <- as.data.frame(NN_for_with_temp_fourier$mean)
print(NN_result)
```
